---
title: "Common barberry - Practice problems - Solutions"
author: "Matt Russell"
date: "11/6/2020"
output: html_document
---

Load the **tidyvers**, **usmap**, **maps**, and **mapproj** packages and read in the `barberry` data set:

```{r}
library(tidyverse)
library(usmap)
library(maps)
library(mapproj)

barberry <- read_csv('barberry.csv')
```

## Practice problems

**1.** In the `barberry` data, which `RecOwner` (e.g., non-profit, state agency, herbarium, etc.) has made the largest number of observations of common barberry?

```{r}
barberry_owner <- barberry %>% 
  group_by(RecOwner) %>% 
  summarize(n = n())

barberry_owner

# Herbarium/Literature/Museums have made the largest number of observations of common barberry (n = 973).
```

**2.** In the `barberry` data, which `Habitat` (e.g., non-profit, state agency, herbarium, etc.) are common barberry most commonly found?

```{r}
barberry_hab <- barberry %>% 
  group_by(Habitat) %>% 
  summarize(n = n())

barberry_hab

# The largest number of observations were for "NA" (n = 1478) and "Other" (n = 155). Roadside/Pipeline/Railroad Habitats had the largest number of observations of common barberry (n = 84).
```

**3.** Make a map of negative reports of common barberry in the US by querying the `Status` variable. First create a new data set using `filter()` with all observations that are negative, e.g., `filter(Status == "Negative")` 

```{r}
barberry_neg <- barberry %>% 
  filter(Status == "Negative")

ggplot() +
 geom_polygon(data = states, aes(x = long, y = lat, group = group), 
              colour = "black", fill = "white" ) + 
    geom_point(data = barberry_neg, aes(x = Longitude, y = Latitude), col = "red") +
  coord_map("conic", lat0 = 30)

# It looks like there are several negative observations in Minnesota and Massachusetts. 
```


