---
title: "Minnesota DNR infested waters"
author: "Matt Russell"
date: "11/6/2020"
output: html_document
---

Load the `tidyverse` package and read in the `waters` data set:

```{r}
library(tidyverse)

waters <- read_csv('infested-waters.csv')
```

## Practice problems - Solutions

**1.** In which year was the greatest number of new infestations observed in Minnesota waters? Make a bar plot that lists the number of AIS by year.

```{r}
ggplot(data = waters, aes(YearInfested)) +
  geom_bar() 
```

**2.** On the `zm` data set, use the `min()` and `max()` functions to calculate the mean, minimum, and maximum values of the number of years between when the zebra mussel was confirmed in a water body to when it became infested (i.e., the `YearsBetween` variable. For example, to calculate the mean number of years between confirmed and infested, you could write `mean(zm$YearsBetween, na.rm = T)`. The `na.rm = T` is used to remove the missing values from the calculation.

```{r}
mean(zm$YearsBetween, na.rm = T)
min(zm$YearsBetween, na.rm = T)
max(zm$YearsBetween, na.rm = T)
```

**3.** Filter the `waters` data to obtain a data set with starry stonewort observations from Beltrami County. How many water bodies are infested?

```{r}
starry <- waters %>% 
  filter(Species == "starry stonewort" & County == "Beltrami")
starry

# Six water bodies are infested with starry stonewort in Beltrami County.
```

**4.** Compile a data set with all Eurasian watermilfoil observations from all counties. Calculate the total number and mean number of years between when Eurasian watermilfoil was confirmed in a water body to when it became infested by each county. (HINT: Eurasian watermilfoil is observed in 39 counties.)

```{r}
milfoil_county <- waters %>% 
  filter(Species == "Eurasian watermilfoil") %>% 
  mutate(YearsBetween = YearInfested - YearConfirmed) %>% 
  group_by(County) %>% 
  summarize(num_waters = n(),
            mean_YearsBetween = mean(YearsBetween, na.rm = T))
milfoil_county
```

**5.** Make a box plot using `geom_boxplot()` that shows the mean number of years between when Eurasian watermilfoil was confirmed in a water body to when it became infested by each county

```{r}
ggplot(data = milfoil_county, aes(mean_YearsBetween)) +
  geom_boxplot() 
```

**6.** Make a bar plot using `geom_bar(stat = "identity")` that shows the number of Eurasian watermilfoil-infested waters in each county. Which county has the most infestations of Eurasian watermilfoil?

```{r}
ggplot(data = milfoil_county, aes(x = num_waters, y = County)) +
  geom_bar(stat = "identity") 

# Hennepin County has the most infestations of Eurasian watermilfoil.
```